Contrastive Optimization -- Preliminaries

### 2024-09-11 paper/main.tex
corrected bug in math

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{gather}
	\max_{\theta} s_c - s_{c'} \equiv \max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c - \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c',i,j}^{\text{HiResCAM}} - b_{c'}
\end{gather}
Ignoring the bias terms, our new optimization objective is:
\begin{gather}
	\max_{\theta} \tilde{\mathcal{\bm{A}}}^{\text{contrastive}} := \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} - \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c',i,j}^{\text{HiResCAM}}
\end{gather}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values preserving spatial information as an activation map instead of flattening the feature map. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect.

For the multiclass case, our setup now considers a set of classes $c$; $c_t$ is our target class, and $c_{t'}$ consists of a sampled non-target class. The definition for multiclass contrastive activation maps now results in a set of activation maps, defined as follows:
\begin{gather}
	\tilde{\mathcal{\bm{A}}}^{\text{contrastive}} := \{\tilde{\mathcal{\bm{A}}}_{c_t}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c_t'}^{\text{HiResCAM}}\}^{|c|-1}_{c_{t'} \in c - c_t}
\end{gather}
This is a generalization of the binary case, as the same formula applied to a binary classification challenge returns a set of contrastive maps containing a single element.

To ensure that the model accurately predicts the target class, each contrastive comparison should yield a positive result. Therefore, the multiclass optimization objective under this paradigm is as follows:
\begin{gather}
	\max_{\theta} \sum^{D_1,D_2}_{j,k}\tilde{\mathcal{\bm{A}}}_{i,j,k}^{\text{contrastive}}\ \forall i \in \mathbb{Z}_{+}(|c| - 1)
\end{gather}

@@

Segmentations maps describe the region over which the object is present. We can leverage this information to define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \text{ where } h_{i,j} := \begin{cases}
		1 & \text{pixel contains subject} \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{background}}_{c,i,j} := \begin{cases}
		\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{c,i,j} & h_{i,j} = 0 \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{foreground}}_{c,i,j} := \begin{cases}
		0 & h_{i,j} = 0 \\
		\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{c,i,j} & \text{otherwise}
	\end{cases} \\
	D_{KL}(P || Q) := \frac{1}{|c|-1}\sum^{|c|-1}_c P \cdot \log \frac{P'}{Q_c} \text{ where } 
	P'_{i,j} := \begin{cases}
		1 & h_{i,j} \neq 0 \\
		P_{i,j} & \text{otherwise}
	\end{cases} \label{d_kl} \\
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_2 \mathcal{\bm{A}}^{\text{contrastive}})\ ||\ \sigma_{\text{softmax}}(\lambda_3 h)) \\
		&+ \frac{\lambda_4}{\sum^{D_1,D_2}_{i,j}\mathbbm{1}(h_{i,j} = 0)} ||\mathcal{\bm{A}}^{\text{background}}||^2_F \\
		&- \frac{\lambda_5}{\sum^{D_1,D_2}_{i,j}\mathbbm{1}(h_{i,j} = 1)} \sum^{|c|-1,D_1,D_2}_{c,i,j} \mathcal{\bm{A}}^{\text{foreground}}_{c,i,j}
	\end{split}
\end{gather}

>>>>

We therefore derive an optimization objective that attempts to maximize the contrast between the two activations maps. For the binary classification case, $(c, c')$ reflects that correct and incorrect class respectively.
\begin{gather}
	\max_{\theta} s_c - s_{c'} \equiv \max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c,i,j}^{\text{HiResCAM}} + b_c - \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}_{c',i,j}^{\text{HiResCAM}} - b_{c'}
\end{gather}
Ignoring the bias terms, our new optimization objective is:
\begin{gather}
	\max_{\theta} \sum^{D_1,D_2}_{i,j} \tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c, c'), i, j} := \tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}}
\end{gather}
This is equivalent to the original ERM optimization objective, with the difference being the resultant values preserving spatial information as an activation map instead of flattening the feature map. If $\tilde{\mathcal{\bm{A}}}_{c}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c'}^{\text{HiResCAM}} > 0$, the classification is correct -- if not, the classification is incorrect.

For the multiclass case, our setup now considers a set of classes $c$; $c_t$ is our target class, and $c_{t'}$ consists of a sampled non-target class. The definition for multiclass contrastive activation maps now results in a set of activation maps, defined as follows:
\begin{gather}
	\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{(c_t, c_{t'})} := \{\tilde{\mathcal{\bm{A}}}_{c_t}^{\text{HiResCAM}} - \tilde{\mathcal{\bm{A}}}_{c_{t'}}^{\text{HiResCAM}}\}^{|c|-1}_{c_{t'} \in c - c_t}
\end{gather}
This is a generalization of the binary case, as the same formula applied to a binary classification challenge returns a set of contrastive maps containing a single element.

To ensure that the model accurately predicts the target class, each contrastive comparison should yield a positive result. Therefore, the multiclass optimization objective under this paradigm is as follows:
\begin{gather}
	\max_{\theta} \sum^{D_1,D_2}_{j,k}\tilde{\mathcal{\bm{A}}}_{(c, c'),j,k}^{\text{contrastive}}\ \forall c' \in \mathbb{Z}_{+}(|c| - 1)
\end{gather}

@@

Segmentations maps describe the region over which the object is present. We can leverage this information to define a constrastive cost function as follows:
\begin{gather}
	\{(x, (h, y))\}^N \in \mathcal{D} \text{ where } h_{i,j} := \begin{cases}
		1 & \text{pixel contains subject} \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{background}}_{c,i,j} := \begin{cases}
		\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{c,i,j} & h_{i,j} = 0 \\
		0 & \text{otherwise}
	\end{cases} \\
	\mathcal{\bm{A}}^{\text{foreground}}_{c,i,j} := \begin{cases}
		0 & h_{i,j} = 0 \\
		\tilde{\mathcal{\bm{A}}}^{\text{contrastive}}_{c,i,j} & \text{otherwise}
	\end{cases} \\
	D_{KL}(P || Q) := \frac{1}{|c|-1}\sum^{|c|-1}_c P \cdot \log \frac{P'}{Q_c} \text{ where } 
	P'_{i,j} := \begin{cases}
		1 & h_{i,j} \neq 0 \\
		P_{i,j} & \text{otherwise}
	\end{cases} \label{d_kl} \\
	\begin{split}
		\mathcal{\bm{J}}(\theta, \mathcal{D}) := &\ \lambda_1 D_{KL}(\sigma_{\text{softmax}}(\lambda_3 h)\ ||\ \sigma_{\text{softmax}}(\lambda_2 \mathcal{\bm{A}}^{\text{contrastive}})) \\
		&+ \frac{\lambda_4}{\sum^{D_1,D_2}_{i,j}\mathbbm{1}(h_{i,j} = 0)} ||\mathcal{\bm{A}}^{\text{background}}||^2_F \\
		&- \frac{\lambda_5}{\sum^{D_1,D_2}_{i,j}\mathbbm{1}(h_{i,j} = 1)} \sum^{|c|-1,D_1,D_2}_{c,i,j} \mathcal{\bm{A}}^{\text{foreground}}_{c,i,j}
	\end{split}
\end{gather}
